market-justice-preferences

### 2024-05-09 paper.tex
update plots and tables analisis

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\includegraphics{input/img/hypotheses.png}

@@

\subsection{}\label{section}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4286}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2551}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1020}}@{}}
\caption{Dependent
variables}\label{tbl-desc-dependientes}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
It is just that in Chile people with higher incomes can have better
pensions than people with low incomes &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1837 (30.6\%)\\
1945 (32.4\%)\\
1622 (27.0\%)\\
608 (10.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6012\\
(95.9\%)\strut
\end{minipage} \\
It is just that in Chile people who can pay have a better education for
their children & \begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1766 (29.7\%)\\
1732 (29.1\%)\\
1704 (28.6\%)\\
750 (12.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5952\\
(94.9\%)\strut
\end{minipage} \\
It is just that in Chile people with higher incomes can access better
health services than people with low incomes &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2254 (38.0\%)\\
1685 (28.4\%)\\
1401 (23.6\%)\\
593 (10.0\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5933\\
(94.6\%)\strut
\end{minipage} \\
Inequality justification index &
\begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 2.2 (0.9)\\
min \textless{} med \textless{} max:\\
1 \textless{} 2 \textless{} 4\\
IQR (CV) : 1.7 (0.4)\strut
\end{minipage} & 13 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
6077\\
(96.9\%)\strut
\end{minipage} \\
\end{longtable}

@@

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-pdf/unnamed-chunk-11-1.pdf}

@@

\caption{Market justice preferences in education, health and pensions by
social and school meritocracy}

@@

Figure 3 shows a series of graphs depicting the association between the
variables of market justice preferences - in education, health, and
pensions - and the variables of meritocratic perception at school
(effort and talent) and in society (effort, talent, and deservingness)
(see conceptual diagram en figure X). On the left, we observe the social
meritocracy diagrams, while on the right the school meritocracy diagrams
are shown. For the three variables of perception of meritocracy in
society the relationship is clear, since the average of market justice
preferences increases the more there is agreement that people are
rewarded for their effort, merit, and talent. This relationship needs to
be clarified in the case of the variables of perception of meritocracy
at school. To the extent that there is more agreement that the
perception of talent is essential for obtaining good grades, the average
of market justice preferences increases, but this relationship is not as
clear as with the variables of meritocracy in society. In addition, the
graph does not show a clear trend in the relationship between the
perception that effort is essential to obtain good grades and market
justice preferences.

@@

Three Cumulative link mixed models were estimated for the ordinal
dependent variables of justice in differential access to pensions,
education, and health according to income. Figure 4 shows the estimation
of this regression model containing all the variables used in the study
for the three dependent variables separately. However, this figure shows
only the effect of meritocracy variables on society and school as
independent variables. Complete models are available in the appendix.

Figure 4 shows that for the three dependent variables of justice in
differential access to pensions, education, and health, the trend is
similar. In the context of school meritocracy, the effects are mixed. As
school talent increases, the justification for differentiated access to
pensions, education, and health increases; on the contrary, as the
school effort variable increases, the justification for differentiated
access to pensions, education, and health decreases, keeping the rest of
the independent variables constant. Regarding the variables of social
meritocracy, the three variables of talent, effort, and deservingness
show that as these increase, the justification for differentiated access
to pensions, education, and health also increases.

@@

Table 4 shows the results of the multilevel estimation for justice
market preferences. For this variable, the intraclass correlation
obtained shows that the variation between schools corresponds to 4\% of
the variation of students' preferences. This means that there is low
variance between schools and therefore limits the possibilities of
finding effects at the aggregate level.

@@

@@

The interaction terms in Table 6 suggest that students' family
background moderates the relationship between their meritocratic
perceptions in Chile and their justification of inequality. The
direction of these effects confirms our initial prediction. Thus, in
line with our Hypothesis 3, the relationship between social effort and
justification of inequality becomes less positive for those students
whose parents achieved university or postgraduate education. That is,
lower-status students (measured by parental education) justify more
inequality when they adhere more strongly to deservingnessmeritocratic
perceptions in Chile. We depict this moderating effect in Figure 1. This
result confirms the enlightenment thesis (see above). The same
moderating effect is observed for students' family cultural capital: the
relationship between effort (but not deservingness nor social talent)
and justification of inequality becomes less positive for students with
more than 25 books at home. Interestingly, family cultural capital also
moderates the relationship between meritocratic beliefs at school (as
measured by students' belief that effort is important to get good
grades) and justification of inequality.

At the school level, we also observe moderating effects of school status
and students' meritocratic beliefs over their justification of
inequality. Thus, in line with our Hypothesis 5, high-status schools
justify less inequality when, on average, their students have a greater
perception of meritocracy in Chile, as measured by the three indicators
we used (i.e., effort, deservingness, and talent). Finally, as we stated
in Hypothesis 7, lowhigh-achieving schools justify more inequality when
their students have, on average, stronger meritocratic beliefs in Chile.
In Figure 25 we depict this moderating effect of school achievement for
the relationship between deservingness and justification of inequality.
We did not observe these moderating effects at the school level for
students' meritocratic beliefs in the school (i.e., the idea that effort
or talent are important to get good grades).

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\includegraphics[width=6.25in,height=\textheight]{input/img/hypotheses.png}

@@

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4286}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2551}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1020}}@{}}
\caption{Dependent
variables}\label{tbl-desc-dependientes}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
It is just that in Chile people with higher incomes can have better
pensions than people with low incomes &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1837 (30.6\%)\\
1945 (32.4\%)\\
1622 (27.0\%)\\
608 (10.1\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
6012\\
(95.9\%)\strut
\end{minipage} \\
It is just that in Chile people who can pay have a better education for
their children & \begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1766 (29.7\%)\\
1732 (29.1\%)\\
1704 (28.6\%)\\
750 (12.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5952\\
(94.9\%)\strut
\end{minipage} \\
It is just that in Chile people with higher incomes can access better
health services than people with low incomes &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly disagree\\
2. Disagree\\
3. Agree\\
4. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
2254 (38.0\%)\\
1685 (28.4\%)\\
1401 (23.6\%)\\
593 (10.0\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
5933\\
(94.6\%)\strut
\end{minipage} \\
Market Justice Preferences & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 2.2 (0.9)\\
min \textless{} med \textless{} max:\\
1 \textless{} 2 \textless{} 4\\
IQR (CV) : 1.7 (0.4)\strut
\end{minipage} & 13 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
6077\\
(96.9\%)\strut
\end{minipage} \\
\end{longtable}

@@

\includegraphics{paper_files/figure-pdf/fig-bivariate-1.pdf}

@@

\caption{\label{fig-bivariate}Market justice preferences in education,
health and pensions by social and school meritocracy}

@@

Figure~\ref{fig-bivariate} shows a series of graphs depicting the
association between the variables of market justice preferences - in
education, health, and pensions - and the variables of meritocratic
perception at school (effort and talent) and in society (effort, talent,
and deservingness) (see conceptual diagram in
Figure~\ref{fig-hypotheses}). On the left, we observe the social
meritocracy diagrams, while on the right the school meritocracy diagrams
are shown. For the three variables of perception of meritocracy in
society the relationship is clear, since the average of market justice
preferences increases the more there is agreement that people are
rewarded for their effort, merit, and talent. This relationship needs to
be clarified in the case of the variables of perception of meritocracy
at school. To the extent that there is more agreement that the
perception of talent is essential for obtaining good grades, the average
of market justice preferences increases, but this relationship is not as
clear as with the variables of meritocracy in society. In addition, the
graph does not show a clear trend in the relationship between the
perception that effort is essential to obtain good grades and market
justice preferences.

@@

Three Cumulative link mixed models were estimated for the ordinal
dependent variables of justice in differential access to pensions,
education, and health according to income. Figure~\ref{fig-odds} shows
the estimation of this regression model containing all the variables
used in the study for the three dependent variables separately. However,
this figure shows only the effect of meritocracy variables on society
and school as independent variables. Complete models are available in
the appendix.

\includegraphics{paper_files/figure-pdf/fig-odds-1.pdf}

\caption{\label{fig-odds}Odds-ratios of justification of education,
pensions and health by social and school meritocracy}

Figure~\ref{fig-odds} shows that for the three dependent variables of
justice in differential access to pensions, education, and health, the
trend is similar. In the context of school meritocracy, the effects are
mixed. As school talent increases, the justification for differentiated
access to pensions, education, and health increases; on the contrary, as
the school effort variable increases, the justification for
differentiated access to pensions, education, and health decreases,
keeping the rest of the independent variables constant. Regarding the
variables of social meritocracy, the three variables of talent, effort,
and deservingness show that as these increase, the justification for
differentiated access to pensions, education, and health also increases.

@@

\begin{table}

\caption{\label{tbl-individual-reg}Individual effects of multilevel
regression models}

\begin{center}
\begin{tabular}{l c c}
\hline
 & Model 1 & Model 2 \\
\hline
Intercept                                    & $1.34^{***}$  & $1.37^{***}$  \\
                                             & $(0.09)$      & $(0.10)$      \\
School talent                                & $0.11^{***}$  & $0.10^{***}$  \\
                                             & $(0.01)$      & $(0.01)$      \\
School effort                                & $-0.12^{***}$ & $-0.12^{***}$ \\
                                             & $(0.02)$      & $(0.02)$      \\
Social talent                                & $0.07^{***}$  & $0.07^{***}$  \\
                                             & $(0.02)$      & $(0.02)$      \\
Social effort                                & $0.08^{***}$  & $0.08^{***}$  \\
                                             & $(0.02)$      & $(0.02)$      \\
Deservingness                                & $0.20^{***}$  & $0.20^{***}$  \\
                                             & $(0.02)$      & $(0.02)$      \\
Parental education (Ref.= 8th grade or less) &               &               \\
                                             &               &               \\
\quad Secondary                              &               & $-0.07$       \\
                                             &               & $(0.05)$      \\
\quad Higher tec.                            &               & $-0.10$       \\
                                             &               & $(0.05)$      \\
\quad University or posgraduate              &               & $-0.02$       \\
                                             &               & $(0.05)$      \\
\quad Missing                                &               & $-0.00$       \\
                                             &               & $(0.05)$      \\
More than 25 books (Ref. Less than 25)       &               & $-0.05$       \\
                                             &               & $(0.03)$      \\
Technology access                            &               & $0.01$        \\
                                             &               & $(0.01)$      \\
\hline
Deviance                                     & $12368.85$    & $12355.60$    \\
Deviance Test (p)                            & $0.00$        & $0.00$        \\
AIC                                          & $12422.86$    & $12468.36$    \\
BIC                                          & $12475.07$    & $12572.78$    \\
Log Likelihood                               & $-6203.43$    & $-6218.18$    \\
Num. obs.                                    & $5047$        & $5047$        \\
Num. groups: mrbd                            & $231$         & $231$         \\
Var: mrbd (Intercept)                        & $0.02$        & $0.02$        \\
Var: Residual                                & $0.66$        & $0.66$        \\
\hline
\multicolumn{3}{l}{\scriptsize{*** p < 0.001; ** p < 0.01; * p < 0.05}}
\end{tabular}
\caption{}
\label{table:coefficients}
\end{center}

\end{table}%

Table~\ref{tbl-individual-reg} shows the results of the multilevel
estimation for justice market preferences. For this variable, the
intraclass correlation obtained shows that the variation between schools
corresponds to 4\% of the variation of students' preferences. This means
that there is low variance between schools and therefore limits the
possibilities of finding effects at the aggregate level.

@@

\begin{table}

\caption{\label{tbl-contextual-reg}Contextual effects of multilevel
regression models}

\begin{center}
\begin{tabular}{l c c}
\hline
 & Model 1 & Model 2 \\
\hline
Intercept                                & $2.29^{***}$  & $1.42^{***}$  \\
                                         & $(0.04)$      & $(0.11)$      \\
Prop. university level at school         & $0.25$        & $0.18$        \\
                                         & $(0.24)$      & $(0.22)$      \\
Administrative dependence (Ref.= Public) &               &               \\
                                         &               &               \\
\quad Subsidized private                 & $0.01$        & $0.00$        \\
                                         & $(0.04)$      & $(0.04)$      \\
\quad Private                            & $-0.07$       & $-0.08$       \\
                                         & $(0.14)$      & $(0.13)$      \\
Socioeconomic level (Ref.= Low)          &               &               \\
                                         &               &               \\
\quad SES Medium low                     & $0.00$        & $0.02$        \\
                                         & $(0.05)$      & $(0.05)$      \\
\quad SES Medium                         & $-0.13$       & $-0.03$       \\
                                         & $(0.07)$      & $(0.06)$      \\
\quad SES Medium high                    & $-0.23^{*}$   & $-0.08$       \\
                                         & $(0.09)$      & $(0.09)$      \\
\quad SES High                           & $0.09$        & $0.25$        \\
                                         & $(0.17)$      & $(0.15)$      \\
Achievement score (Ref.= Low)            &               &               \\
                                         &               &               \\
\quad Simce Medium                       & $-0.09^{*}$   & $-0.10^{**}$  \\
                                         & $(0.04)$      & $(0.04)$      \\
\quad Simce High                         & $-0.27^{***}$ & $-0.24^{***}$ \\
                                         & $(0.05)$      & $(0.04)$      \\
\hline
Deviance                                 & $12889.80$    & $12306.45$    \\
Deviance Test (p)                        & $0.00$        & $0.00$        \\
AIC                                      & $12954.69$    & $12473.31$    \\
BIC                                      & $13033.01$    & $12636.48$    \\
Log Likelihood                           & $-6465.35$    & $-6211.66$    \\
Num. obs.                                & $5047$        & $5047$        \\
Num. groups: mrbd                        & $231$         & $231$         \\
Var: mrbd (Intercept)                    & $0.02$        & $0.01$        \\
Var: Residual                            & $0.74$        & $0.66$        \\
\hline
\multicolumn{3}{l}{\scriptsize{*** p < 0.001; ** p < 0.01; * p < 0.05}}
\end{tabular}
\caption{}
\label{table:coefficients}
\end{center}

\end{table}%

@@

\begin{table}

\caption{\label{tbl-interact}Interactions effects}

[H]
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{5}{c}{\textbf{Market Justice Preferences}} \\
\cmidrule(l{3pt}r{3pt}){2-6}
Variable & School talent & School effort & Social talent & Social effort & Deservingness\\
\midrule
\textbf{University or Postgraduate} & -0.034 & -0.012 & -0.073. & -0.079* & -0.118**\\
\cmidrule{1-6}
\textbf{More than 25 books} & -0.015 & -0.046 & -0.029 & -0.081** & -0.035\\
\cmidrule{1-6}
\textbf{School SES high} & 0.004 & 0.001 & -0.057. & -0.063* & -0.08**\\
\cmidrule{1-6}
\textbf{Prop. university level} & 0.008 & -0.021 & -0.123 & -0.171* & -0.162.\\
\cmidrule{1-6}
\textbf{Simce Medium} & 0.001 & 0.032 & -0.037 & -0.041 & -0.082*\\
\cmidrule{1-6}
\textbf{Simce High} & -0.016 & 0.008 & -0.043 & -0.063. & -0.087*\\
\bottomrule
\multicolumn{6}{l}{\rule{0pt}{1em}\textit{Note: } ***p < 0.001, **p < 0.01, *p < 0.05}\\
\end{tabular}}

\end{table}%

The interaction terms in Table~\ref{tbl-interact} suggest that students'
family background moderates the relationship between their meritocratic
perceptions in Chile and their justification of inequality. The
direction of these effects confirms our initial prediction. Thus, in
line with our Hypothesis 3, the relationship between social effort and
justification of inequality becomes less positive for those students
whose parents achieved university or postgraduate education. That is,
lower-status students (measured by parental education) justify more
inequality when they adhere more strongly to deservingnessmeritocratic
perceptions in Chile. This result confirms the enlightenment thesis (see
above). The same moderating effect is observed for students' family
cultural capital: the relationship between effort (but not deservingness
nor social talent) and justification of inequality becomes less positive
for students with more than 25 books at home. Interestingly, family
cultural capital also moderates the relationship between meritocratic
beliefs at school (as measured by students' belief that effort is
important to get good grades) and justification of inequality.

At the school level, we also observe moderating effects of school status
and students' meritocratic beliefs over their justification of
inequality. Thus, in line with our Hypothesis 5, high-status schools
justify less inequality when, on average, their students have a greater
perception of meritocracy in Chile, as measured by the three indicators
we used (i.e., effort, deservingness, and talent). Finally, as we stated
in Hypothesis 7, low-achieving schools justify more inequality when
their students have, on average, stronger meritocratic beliefs in Chile.
In Figure~\ref{fig-interaction} we depict this moderating effect of
school achievement for the relationship between deservingness and
justification of inequality. We did not observe these moderating effects
at the school level for students' meritocratic beliefs in the school
(i.e., the idea that effort or talent are important to get good grades).

\includegraphics{paper_files/figure-pdf/fig-interaction-1.pdf}

\caption{\label{fig-interaction}Interaction between deservingness and
justification of inequality by school SIMCE achievement}
